---
title: "Homework 2"
subtitle: "The Probability Function"
author: "WRITE YOUR NAME HERE"
output: 
    pdf_document:
        highlight: null
        pandoc_args: ["--syntax-highlighting=tango"]
header-includes:
    - \usepackage{amsmath, amssymb}
    - \usepackage{setspace} 
---
\onehalfspacing

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)  #R package 
```

* * * 

### Instructions

- This homework is due in Gradescope on Wednesday Jan 28 by midnight PST. There is a 15 minute grace period and submissions made during this time will not be marked as late. Any work submitted past this period is considered late.  

- Please answer the following questions in the order in which they are posed. \textbf{Be sure to support your work and show your steps. All calculations must be done within your document and shown with the answer printed.}

- Don't forget to (i) make a local copy this document for your work and to (ii) knit the document frequently to make sure there are no compilation errors. 

- When you are done, download the PDF file as instructed in section and submit it in Gradescope. \textbf{Do not forget to tag your answers to every sub-part of every question. We will take off 10\% for every part that is not tagged}

- \textbf{Please read the acceptable policies for this class in CANVAS. In particular, you are NOT ALLOWED to consult external sources - AI, tutors, tutoring sites, internet - to complete any part of this work. Do not use notation or definitions or results we have not covered in this class. We will deduct points for this. }

* * *


### Exercises 

<!--- This is a hint for Problem 1 part b. It won't show up in your PDF. You can delete it after reading if you wish.

Write $E = E_4  \cup E_5 \cup E_6$ where $E_4$ denotes the event there are 4 heads, etc. Figure out how many elements there are in $E_4$, $E_5$ and $E_6$.
--->

1. (Heads or Tails) Suppose 6 \emph{fair} coins are tossed and the outcome (head or tail) is recorded for each.  
   
a. The sample space $S$ corresponding to this "experiment" consists of all possible sequences of heads and tails that result from tossing 6 coins. How many elements are in $S$? Calculate the answer in a code chunk, save it to a variable called `num_S` and then print the value.  
   
b. Let $E$ denote the event that there are more heads than tails. How many elements are in $E$? Calculate this number in a code chunk, save your answer in a variable called `num_E` and print the value.  

   
c. What is $P(E)$? Calculate it in a code chunk and save it to variable called `prob_E` and print the value. Be sure to cite any rules you are using to make your theoretical probability calculation. 
   

<!--- This is a hint for Problem 2 part e. It won't show up in your PDF. You can delete it after reading if you wish.

Think in terms of the fraction of the population which is tagged and the fraction of the observed sample which is tagged to understand the pattern.
--->


2. (Tagging goldfish) To estimate the number $N$ of goldfish in a pond, $r = 25$ fish were caught, tagged and released. Later, a second sample of $n = 20$ fish were caught and 5 fish in this sample were noted to be tagged.  
 
a.  How many possible samples of size $n = 20$ can be formed from the $N$ fish in the pond?  Leave your answer in terms of a binomial coefficient since you cannot calculate it because you don't know $N$. 
    
b. The event $E$ contains all the samples which have 5 tagged and 15 untagged fish. How many elements are in the event $E$? Leave your answer in terms of $N$. 
   
c. Assuming each possible sample is equally likely, give an expression for $P(E)$. (Leave your answer in terms of $N$)

d.  In this part, we will examine visually how $P(E)$ varies as a function of $N$.  Fill in the blanks in the R code provided and run it in R to create the plot. **You should remove the `eval = FALSE` chunk option after filling in the blanks.**


    ```{r fish, eval=FALSE}

    fishes <- tibble( #enhanced data frame
              N = 40:200,    #possible values for N: 40,41, ...,200
              prob = ___ ) ) #P(E) in terms of N

    ggplot(data = fishes, 
           mapping = aes(x = ___,    
                         y = ___) ) +
           geom_line() +
             labs( title = " ___" ,
                  x = "___", 
                  y = "___")
    ```


e. If all went well, your graph should be unimodal with a peak at $N=100$. Just conceptually, explain what this pattern tells us about $P(E)$ and why it makes sense. Be sure to demonstrate your understanding of the location of the peak, as well as the increasing pattern before the peak and decreasing pattern after.

<!--- 
This is a hint for writing up your work for Problem 3. It won't show up in your PDF. You can delete it after reading if you wish.

Take a look at my explanations for Example 3.1 in the slide deck for chapter 3. We will deduct credit if you do not justify your work or show steps in a similar manner. 
--->

3. (Just acting) An actress playing the lead role in a film has a stunt double. According to the script, the actress appears in 40% of the film’s scenes, her double appears in 30%, and the two of them are together 5% of the time. Let $L$ denote the event that the lead is in a scene and $D$ denotes that the double is in a scene. Write each of the following in terms of $L$ and $D$ and then find the probability that it occurs.

a. only the stunt double appears in a scene? 
   
b. neither the lead actress nor the double appears in a scene? 


<!--- 
This is a hint for writing up your work for Problem 4. It won't show up in your PDF. You can delete it after reading if you wish.

Refer to Bonferroni's equality from Chapter 3 slide deck and also the subset inequality from problem section 2. You will need them.
--->

4.  (Bounds) For two events $A$ and $B$ with $P(A)=0.5$ and $P(B)=0.8$, what are the largest and smallest possible values for $P(A \cap B)$? Be sure to show your work and explain the steps. 



