---
title: "Problem Section 4"
subtitle:  "Discrete Distributions"
graphics: yes
output: 
        pdf_document
header-includes:
    - \usepackage{amsmath, amssymb}
    - \usepackage{framed}\definecolor{shadecolor}{rgb}{0.949,0.949,0.949}
---

```{r setup, include=FALSE}
#use this code chunk to set global options
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
```


* * * 
\begin{shaded}

\textbf{Learning Outcomes}

The problems are designed to build conceptual understanding and problem-solving skills. The emphasis is on learning to find, evaluate and build confidence.
The specific tasks include: 

    - Identifying the range of discrete and continuous random variables
   
   - Deriving the PMF and CDF of a discrete random variable
   
   - Recognizing properties of CDFs
   
   - Back up and support work with relevant explanations
   

\end{shaded}

* * *

### Exercises 

1. For each situation, give the range of the random variable $X$. Then state whether the random variable is discrete or continuous.

a. $X$ is the number of employees among 12 recently hired that will still be with the company in another year.
    	
    The range is $\{0,1,2,\dots, 12\}$. $X$ is discrete since it takes finitely many values.
		
b. In a busy 24 hour copy center, minor malfunctions of the copy machine occur randomly. $X$ denotes the time (in hours) until the first minor malfunction occurs.
	 
	 The range is $[0, \infty)$. Here, $X$ can take any value in an interval and can (in theory) be measured with infinite precision. It is a continuous random variable even though it is being rounded to hours.
	 
2. 2. Suppose you toss a fair coin 3 times. The sample space for this experiment is shown below:
$$S = \{ HHH, HHT, HTH, HTT, TTT,TTH, THT, THH \}.$$

a. Let $X$ denote the difference between number of heads and number of tails. Write the PMF of $X$ in a tabular format.   

    \begin{tabular}{|c|c|c|c|c|} \hline
    $x$ & -3 & -1 & 1 & 3 \\ \hline
    $f(x)$ & 1/8 & 3/8 & 3/8 & 1/8 \\ \hline
    \end{tabular}

b. Draw the probability histogram of $X$ in R. Comment on the appearance of the distribution such as shape, modality.

    ```{r prob-hist}
    # be sure to add library(tidyverse) in setup chunk. 
    #Do not add it here, we usually add packages in the setup code chunk
    #This code chunk has chunk options set so that the code is not printed,
    #messages while loading packages are not displayed, etc.

    ht_df <- tibble(
      x = c(-3,-1,1,3),
      f = c(1/8,3/8,3/8,1/8))

    ggplot(data=ht_df, 
         mapping = aes(x = x, y = f) )+
      geom_col(width = 1, alpha = 0.5,         #don't forget to set that width
           fill = "gold", color = "purple") +
      labs(title = "Probability Histogram of X",
            y = "probability",
            x = "X = x") #+
          #scale_x_continuous(breaks = c(-3,-1,1,3))

    ```



3. Which of the following is a valid CDF? (choose one; also point out what's wrong with the ones you didn't choose)

Hint: make a sketch 

a. 
\begin{align*} 
        F(x) &= \left\{ \begin{array}{cc} 
                 0 & x < 0\\
                 \frac{1}{4} & 0 \leq x \leq 3 \\
                 \frac{3}{4} & 3 <  x < 5 \\
                 1 & 5 \leq x \end{array} \right. \end{align*}

    This is not a valid CDF because it is not right continuous at $x=3$. The function should take the value at the top of the jump.
    
b.
\begin{align*} 
        F(x) &= \left\{ \begin{array}{cc} 
                 0 & x < 0\\
                 \frac{1}{4} & 0 \leq x < 3 \\ 
                 \frac{3}{4} & 3 \leq x < 5 \\ 
                 1 & 5 \leq x \end{array} \right. \end{align*}

    This is a valid CDF because it is non-decreasing, right continuous, $F(\infty) = 1$ and $F(-\infty) = 0$.

c. 
\begin{align*} 
      F(x) &= \left\{ \begin{array}{cc} 
                 0 & x < 0\\
                 \frac{3}{4} & 0 \leq x < 3 \\ 
                 \frac{1}{4} & 3 \leq x < 5 \\ 
                 1 & 5 \leq x \end{array} \right. \end{align*}
   
This is not a valid CDF since it decreases at $x=3$. A CDF is always non-decreasing.


d.
\begin{align*} 
       F(x) &= \left\{ \begin{array}{cc} 
                 0 & x < 0\\
                 \frac{1}{4} & 0 \leq x < 3 \\ 
                 \frac{3}{4} & 3 \leq x < 5 \\ 
                 \frac{7}{8} & 5 \leq x \end{array}   
                 \right.                      
\end{align*}              

This is not a valid CDF since it does not converge to 1 as $x \rightarrow \infty$.

4. Write the PMF for the CDF that you identified as valid in problem 3. Also calculate the following probabilities using both the CDF and also verify using the PMF.  

    The PMF is
    \begin{align*}
    f(x) &= \left\{ \begin{array}{cc}
                      \frac{1}{4} & x = 0\\
                      \frac{1}{2} & x = 3 \\
                      \frac{1}{4} & x = 5
                      \end{array} \right.
    \end{align*}
    
    The following are probabilities calculated from the CDF in part b.
    
a. $P(X > 2)$ 

    This is simply $1 - P(X \leq 2) = 1 - F(2) = 1-\frac{1}{4} = \frac{3}{4}$. 

b. $P(0.5 < X < 3.5)$

    We know $P(0.5 < X < 3.5) = P(X < 3.5) - P(X \leq 0.5)$. The first probability is simply the left hand limit of $F(x)$ at $x = 3.5$. Therefore $P(X < 3.5)= \frac{3}{4}$.
    The second probability is $F(0.5) = \frac{1}{4}$. Therefore $P(0.5 < X < 3.5) = 0.5$.
    
    

