---
title: "Problem Section 3"
subtitle:  "Conditional Probability & Independence"
graphics: yes
output: 
    pdf_document:
        highlight: null
        pandoc_args: ["--syntax-highlighting=tango"]
header-includes:
    - \usepackage{amsmath, amssymb}
    - \usepackage{framed}\definecolor{shadecolor}{rgb}{0.949,0.949,0.949}
---

```{r setup, include=FALSE}
#use this code chunk to set global options
knitr::opts_chunk$set(echo = TRUE)

###attach R packages which may be needed
library(openintro)    #contains R function for drawing probability tree

```


* * * 
\begin{shaded}

\textbf{Learning Outcomes}

The problems are designed to build conceptual understanding and problem-solving skills. The emphasis is on learning to find, evaluate and build confidence.
The specific tasks include: 

   - Calculate conditional probabilities
   
   - Apply Bayes' rule
   
   - Calculate probabilities for independent events
  
   - Back up and support work with relevant explanations


\end{shaded}

* * *

### Exercises 

1. Suppose a card is drawn randomly from a 52 card standard deck consisting of 13 cards (in denominations from 2 to 10, Jack, King, Queen, Ace) from each of four suits ($\heartsuit$, $\diamondsuit$, $\clubsuit$, $\spadesuit$). 

a. Find the probability that the card is a club, given that the card is a King.

b. Are the events independent? Why or why not?

2. One common application of conditional probabilities is in calculating inverse probabilities. For example, consider the following tree diagram of probabilities relating a screening test for a disease with the actual presence of disease. The first (primary) branch of the tree gives \emph{unconditional} probabilities of disease being present or absent, while the secondary branches state conditional probabilities of the test coming out positive (or negative) given the disease status.


    ```{r treediagram}

    treeDiag(main = c("Disease status", "Test result"),
         p1 = c(0.001, 0.999),
         p2 = list(c(0.98, "?"), c(0.01, "?")),
         out1 = c("Present", "Absent"),
         out2 = c("Positive", "Negative"),
         showSol = FALSE,
         showWork = FALSE,
         solwd = 0.5,
         digits = 6)

    ```

a. Fill in the values in the tree diagram indicated by a "?". 

b. Suppose the test is positive. What is the probability that the person actually has disease? 

3. A weather satellite is sending a binary code of 0s or 1s describing a developing tropical storm. Channel noise introduces a certain amount of transmission errors. Suppose that 70% of the message being relayed is 0s, and there is an 80% chance of a given 0 or 1 being received properly. If a 1 is received, what is the probability that a 0 was sent?

   Make a probability tree and find the desired probability.
   
4. Suppose that string of tree lights you just bought has twenty-four bulbs
wired in series. If each bulb has a 99.9% chance of “working” when the current
is applied, what is the probability that the string itself will not work? You may assume that bulb failures are independent. 

    In addition to writing up your steps, create a new code chunk below and do your calculations there. Be sure to print the result.